---
title: "ST684: Assignment 1"
author: "Nishant Kumar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 4,fig.height = 4)
```


```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(GGally))

```

###question 2


```{r}
data(alcohol, package="robustbase")
alcohol <- alcohol[,-7]

```


##(a) Scale the data using the default options. Construct the euclidean distance matrix ofthe cases. Cluster the cases, using average linkage. Draw the dendrogram.
```{r}
view(alcohol)
x1<-scale(alcohol,center = TRUE,scale = TRUE)
d1<-as.matrix(dist(x1,method="euclidean"))
d1
h1<-hclust(as.dist(d1),"average")
d6<-as.dendrogram(h1)
plot(d6)
```

(b) Examine the 4-cluster solution. How many cases are in each cluster? Summarise the partitions with sumPartition (in h1code.R) Be sure to use the scaled data. Comment on the chemical composition of the four clusters.
```{r}
d7<-colord5<-color_branches(d6,k=4,col=c(1:4))
plot(d7)
sumPartition(x1,cutree(h1,4))

```

(c) Verify your findings by drawing a parallel coordinate plot of the data coloured by the clusters. Choose a suitable scaling.
```{r}
p1<-cutree(h1, 4)
p1
ax<-data.frame(x1,p1)
head(ax)
ggparcoord(ax,col=1:6,groupColumn = "p1",scale = "globalminmax")

```
(d) Use the kmeans algorithm to find another 4-cluster grouping. Use the scaled data.How many cases are in each cluster?
```{r}
km2<-kmeans(x1,4,nstart=10)
km2
ax1<-data.frame(ax,km2$cluster)
head(ax1)
```
(e) Construct a stars plot which shows the data and clustering obtained from kmeans.Arrange the stars by cluster. (Advanced: arrange the stars by size).

```{r}
help(stars)
stars(ax1[h1$order,],nrow=4,col.stars=km$cluster[h1$order]+1)

stars(ax1[h1$order,], nrow=4, col.stars=cutree(h,4)[h1$order]+1)
####looking for answers
```

3. Eight online shoppers buy 8, 11, 7 , 6, 5, 6 , 7, 8 pairs of socks. The same eight shoppers buy 0,0,0,0,1,1,1,1 computers.

```{r}
x<-c(8,11,7,6,5,6,7,8)
y<-c(0,0,0,0,1,1,1,1)
xy<-data.frame(x,y)
xy
km3<-kmeans(xy,2,nstart=10)
km3
```
(b) If both variables are scaled to unit standard deviation, what will kmeans with k = 2 give you?

```{r}
help(scale)
x1<-scale(xy,center=FALSE,scale=apply(xy,2,sd,na.rm=TRUE))
kmean<-kmeans(xy,2,nstart=10)
kmean
```
(c) Suppose socks cost 2 euro and the computer is 2000 euro. What if you clustered the amount spent by each customer using kmeans with k = 2, with no scaling?
```{r}
xy$x<-xy$x*2
xy$x
xy$y<-xy$y*2000
xy$y
kme<-kmeans(xy,2,nstart=10)
kme

```

#### Question 4



```{r} 
w <-read.csv("/users/students/19251001/R_workings/worldhappiness2019.csv")
names(w)<- c( "Rank", "Country", "Score", "GDP", "Family", "Life", "Freedom", "Generosity", "Trust")
rownames(w)<-w$Country
wscores<-w[,-(1:3)]
```

(a)Calculate the correlation matrix of the scores. Which pair of variables have the highest correlation? Make a scatterplot of the scores for these two variables. Find the name of the outlying country on this graph.

```{r}

cor(wscores,method="spearman")
#GDP and life has highest correlation.
w1<-data.frame(wscores,w$Country)
p2<-ggplot(w1,aes(x=w1$GDP,w1$Life))+geom_point()
p2+geom_text(aes(label=w$Country))
#swaziland is the outlying country on the graph
```
The pair of variables with the highest correlation are ??.
The outlying country on this graph is ??.

(b)Construct the euclidean distance matrix wscores (no standardisation). Use it to cluster the countries, using average linkage. Draw the dendrogram. Are there any outlier countries?
```{r}
suppressPackageStartupMessages(library(dendextend))
d1<-as.matrix(dist(wscores,method="euclidean"))
h<-hclust(as.dist(d1),"average")
d4<-as.dendrogram(h)
d3<-colord5<-color_branches(d4,k=3,col=c(1:3))
plot(d3)
#Central African Republic is the outlier country

```

(c)How many countries are in the clusters of the three cluster solution? Which cluster does Ireland belong to? India? Which cluster has the countries with the highest Family score? (use sumPartition). Which cluster has the highest scores?
```{r}

p<-cutree(h, 3)
p
source("/users/students/19251001/R_workings/semester2/h1code.R")
sumPartition(wscores,cutree(h,3))
#cluster sizes are 111 in 1,44 in cluster 2 ,1 in cluster 3
#total countries in three cluster solution is 156
#Ireland belongs to cluster 1
#India belongs to cluster 2
#Cluster 1 has the highest Family score.
#Cluster 1 has the highest  scores overall.

```

(d)Make a parallel coordinate plot of the scores coloured by the clusters. Choose a suitable scaling. What is unusual about the lowest-scoring cluster from this graph?
```{r}

w2<-data.frame(w1,p)
w2$p
ggparcoord(w2,columns=1:6, groupColumn="p",scale="globalminmax")
#The value of Family is almost equal to zero and GDP is also near to the zero #for the third cluster but freedom and generosity fairly better to the other #countires with good GDP and family scores 

```

The lowest scoring cluster is unusual becuase ???

(e)Cluster the countries using kmeans (no standardisation), use nstart=10, centers=3,and a seed of 123. Are there any outlier countries?
```{r}
set.seed(123)
km <- kmeans(wscores,3,nstart= 10)
tab<-table(w1$w.Country,km$cluster)
tab
#There are no outlier countries

```
(f)Make a parallel coordinate plot of the scores coloured by the kmeans clusters. Choose a suitable scaling. From the graph, which cluster has the countries with the highest Family score? Which cluster has the highest scores overall?
```{r}
km1<-as.factor(km$cluster)
w3<-data.frame(w1,km1)
head(w3)
ggparcoord(w3,columns=1:6, groupColumn="km1",scale="globalminmax")

#cluster 1 has highest family scores
sumPartition(wscores,km$cluster)
#cluster 1 has highest overall scores.

```


(g)For the kmeans clustering result, make a boxplot of Score (in w) for the three clusters.Comment on how the Scores of the clusters compare.
```{r}

boxplot(w$Score~w3$km1)
#The score of the cluster 1 has highest value followed by the cluster 2 value 
#And then the cluster 3 values as the lowest.
#The mean value of the cluster 1 is in range 6 to 7.
#The mean value of the cluster 2 is in range  5 and 6.
#The mean value of the cluster 3 is in range of 4 to 5.

```


